---
title: "Dennery lab scRNAseq"
author: "Joselynn"
date: "7/14/2020"
output: html_document
---

#add depp1 etc genes back in
#re-order the groups on the x-axis 'Macrophage", "AT2, AT1, Mesenchymal, Endo, other immume

#rerun stats with just findmarkers to  make a specific pairwise comparison to sd7 and o2d7
#in addition to like supplemental table 1, include the actual file name as like a comment
#add some text to the results section with the distributions of the macrophage groups

#push updates to git
#add some AT2 marker gene figures -- do the macrophage seurat clusters have different AT2 markers? could just look at overlaps  

  
```{r "Setup"}
#install.packages('rstatix')
.libPaths(c("/usr/local/lib/R/site-library","/usr/local/lib/R/library"))
library(Seurat)
library(magrittr)
set.seed(61)
options(future.globals.maxSize = 4000 * 1024^3)
future::plan("multisession", workers = 4)
load('/oscar/data/cbc/Dennery_scRNAseq/dennery_singlecell_reanalysis/analysis.RData')#
#load('/users/jwalla12/dennery_singlecell_reanalysis/analysis.RData')

```

#Find markers distinguishing different groups within SD7 (where groups are a combo of cell type annotation and seurat cluster):
```{r}
library(magrittr)
Idents(all_data.combined) <- 'grouped_seurat'
grouped_seurat_markers <- FindAllMarkers( 
  subset(all_data.combined, subset = orig.ident == 'SD7'), only.pos = T, min.pct = .2)

#sd7_sub <- subset(all_data.combined, subset = orig.ident == 'SD7')
#Idents(sd7_sub) <- 'grouped_seurat'
#sd7_grouped_markers <- FindAllMarkers(sd7_sub, only.pos = T, min.pct = .2)
```

#Find markers distinguishing the atlas predicted groups
```{r}
Idents(all_data.combined) <- 'atlas_predicted_clusters_short'
atlas_type_markers <- FindAllMarkers(all_data.combined, only.pos = T, min.pct = .2)

clustering_atlas_summary <- table(all_data.combined@meta.data$seurat_clusters, all_data.combined@meta.data$atlas_predicted_clusters_short) %>% rbind()  %>% data.frame() 

clustering_atlas_summary$seurat_cluster <- rownames(clustering_atlas_summary)

macrophage_clustering_atlas_summary <- clustering_atlas_summary %>% dplyr::select('seurat_cluster', 'Alveolar.macrophage_Ear2.high', 'Alveolar.macrophage_Pclaf.high', 'Interstitial.macrophage')

keep <- rowSums(macrophage_clustering_atlas_summary[,c(2:4)] > 0) >= 1
macrophage_clustering_atlas_summary <- macrophage_clustering_atlas_summary[keep,]

#write.table(macrophage_clustering_atlas_summary, file = '/users/jwalla12/dennery_singlecell_reanalysis/data/macrophage_clustering_atlas_summary.csv', row.names = F)

#Ear2, Pclaf, Mrc1, Itgax, Siglecf, Siglec1, Cd14, Ly6c1, Apoe, Ccr5, Mafb, Sell
ram_and_recam_genes <- c('Mrc1', 'Itgax', 'Siglecf', 'Siglec1', 'Cd14', 'Ly6c1', 'Apoe', 'Ccr5', 'Mafb', 'Sell')
ram_genes <- c('Mrc1', 'Itgax', 'Siglecf', 'Siglec1')
recam_genes <- c('Cd14', 'Ly6c1', 'Apoe', 'Ccr5', 'Mafb', 'Sell')

Alveolar_macrophage_Ear2_high_markers <- atlas_type_markers %>% dplyr::filter(cluster == "Alveolar macrophage_Ear2 high") %>% dplyr::filter(p_val_adj <= 0.05)

Alveolar_macrophage_Pclaf_high_markers <- atlas_type_markers %>% dplyr::filter(cluster == "Alveolar macrophage_Pclaf high") %>% dplyr::filter(p_val_adj <= 0.05)

Interstitial_macrophage_markers <- atlas_type_markers %>% dplyr::filter(cluster == "Interstitial macrophage") %>% dplyr::filter(p_val_adj <= 0.05)
```

```{r}
Alveolar_macrophage_Ear2_high_markers_sorted <- Alveolar_macrophage_Ear2_high_markers %>% dplyr::arrange(desc(avg_log2FC)) %>% dplyr::filter(pct.1 >= .7) 

Alveolar_macrophage_Pclaf_high_markers_sorted <- Alveolar_macrophage_Pclaf_high_markers %>% dplyr::arrange(desc(avg_log2FC)) %>% dplyr::filter(pct.1 >= .7)

Interstitial_macrophage_markers_sorted <- Interstitial_macrophage_markers %>% dplyr::arrange(desc(avg_log2FC)) %>% dplyr::filter(pct.1 >= .7)
```

```{r}
IM_SD7_Only <- subset(all_data.combined, subset = (orig.ident == 'SD7' &  
                                    atlas_predicted_clusters_short == 'Interstitial macrophage'))

IM_SD7_Only_clusters <- data.frame(as.matrix(table(IM_SD7_Only@meta.data$seurat_clusters)))
colnames(IM_SD7_Only_clusters) <- c('cell_count')
IM_SD7_Only_clusters$seurat_cluster <- rownames(IM_SD7_Only_clusters)
IM_SD7_Only_clusters$orig.ident <- 'SD7'

ggplot2::ggplot(IM_SD7_Only_clusters, ggplot2::aes(x = orig.ident, y = seurat_cluster)) +
  ggplot2::geom_tile(ggplot2::aes(fill = cell_count)) +
  ggplot2::geom_text(ggplot2::aes(label = cell_count)) +
  ggplot2::scale_fill_gradient(low = "white", high = "tomato") +
  ggplot2::theme(panel.grid = ggplot2::element_blank(),
        axis.text.x = ggplot2::element_text(angle=90),
        axis.ticks = ggplot2::element_blank()) +
  ggplot2::coord_equal()




conc_plot <- ggplot(conc_plot_data, aes(x, y, fill = value)) +
  geom_tile(colour = 'black') + 
  theme(panel.grid = element_blank(),
        axis.title.y = element_blank(),
        axis.title.x = element_blank(),
        axis.text.x = element_blank(),
        axis.ticks = element_blank()) +
  scale_fill_gradientn(
    colors = c("white", "darkred"),
    limits = range(conc_plot_data$value)) +
  theme(panel.grid = element_blank(),
        axis.text.x=element_text(angle=90),
        axis.ticks = element_blank())

```

```{r}
ear2_dp <- DotPlot(subset(all_data.combined, subset = (seurat_clusters == '13' | 
                                    seurat_clusters == '16' | 
                                    seurat_clusters == '17' | 
                                    seurat_clusters == '20' | 
                                    seurat_clusters == '23' | 
                                    seurat_clusters == '25' | 
                                    seurat_clusters == '7')),
        features = Alveolar_macrophage_Ear2_high_markers_sorted$gene[1:5], group.by = 'orig.ident' ) + 
  ggplot2::scale_colour_gradient2(low = "#c6dbef", mid = "#6baed6", high = "#084594") + 
  ggplot2::theme(axis.text.x = ggplot2::element_text(angle = 45, vjust = 1, hjust=1)) + 
  ggplot2::labs(title = "Ear2 High Macrophage Markers")


pclaf_dp <- DotPlot(subset(all_data.combined, subset = (seurat_clusters == '13' | 
                                    seurat_clusters == '16' | 
                                    seurat_clusters == '17' | 
                                    seurat_clusters == '20' | 
                                    seurat_clusters == '23' | 
                                    seurat_clusters == '25' | 
                                    seurat_clusters == '7')),
        features = Alveolar_macrophage_Pclaf_high_markers_sorted$gene[1:5], group.by = 'orig.ident' ) + 
  ggplot2::scale_colour_gradient2(low = "#e5f5e0", mid = "#74c476", high = "#005a32") + 
  ggplot2::theme(axis.text.x = ggplot2::element_text(angle = 45, vjust = 1, hjust=1))  + 
  ggplot2::labs(title = "Pclaf High Macrophage Markers")

interst_dp <- DotPlot(subset(all_data.combined, subset = (seurat_clusters == '13' | 
                                    seurat_clusters == '16' | 
                                    seurat_clusters == '17' | 
                                    seurat_clusters == '20' | 
                                    seurat_clusters == '23' | 
                                    seurat_clusters == '25' | 
                                    seurat_clusters == '7')),
        features = Interstitial_macrophage_markers_sorted$gene[1:5], group.by = 'orig.ident' ) + 
  ggplot2::scale_colour_gradient2(low = "#fcbba1", mid = "#fb6a4a", high = "#99000d") + 
  ggplot2::theme(axis.text.x = ggplot2::element_text(angle = 45, vjust = 1, hjust=1))  + 
  ggplot2::labs(title = "Interstitial Macrophage Markers")

ear2_dp +pclaf_dp + interst_dp
ggplot2::ggsave(here::here('data/all_idents_macro_subtypes_top5_dotplot.png'),
       device = 'png', width = 18, height = 4 )
#Saving 16.2 x 9.66 in image


```

```{r}
ram_and_recam_genes %in% Alveolar_macrophage_Ear2_high_markers$gene

DotPlot(all_data.combined, features = ram_and_recam_genes, split.by = 'atlas_predicted_clusters_short')
all_data.combined@meta.data$macros_or_not <- all_data.combined@meta.data$atlas_predicted_clusters_short

celltypes <- all_data.combined@meta.data$macros_or_not %>% unique() 
notmacros_celltypes <- celltypes[!c(celltypes %in% c("Alveolar macrophage_Ear2 high", "Alveolar macrophage_Pclaf high" , "Interstitial macrophage"))]

```

```{r}
all_data.combined@meta.data$macros_or_not <- stringr::str_replace_all(all_data.combined@meta.data$macros_or_not ,
                         notmacros_celltypes[1],
                         'not_macrophage')

all_data.combined@meta.data$macros_or_not <- stringr::str_replace_all(all_data.combined@meta.data$macros_or_not ,
                         notmacros_celltypes[2],
                         'not_macrophage')

all_data.combined@meta.data$macros_or_not <- stringr::str_replace_all(all_data.combined@meta.data$macros_or_not ,
                         notmacros_celltypes[3],
                         'not_macrophage')

all_data.combined@meta.data$macros_or_not <- stringr::str_replace_all(all_data.combined@meta.data$macros_or_not ,
                         notmacros_celltypes[4],
                         'not_macrophage')

all_data.combined@meta.data$macros_or_not <- stringr::str_replace_all(all_data.combined@meta.data$macros_or_not ,
                         notmacros_celltypes[5],
                         'not_macrophage')

all_data.combined@meta.data$macros_or_not <- stringr::str_replace_all(all_data.combined@meta.data$macros_or_not ,
                         notmacros_celltypes[6],
                         'not_macrophage')

all_data.combined@meta.data$macros_or_not <- stringr::str_replace_all(all_data.combined@meta.data$macros_or_not ,
                         notmacros_celltypes[7],
                         'not_macrophage')

all_data.combined@meta.data$macros_or_not <- stringr::str_replace_all(all_data.combined@meta.data$macros_or_not ,
                         notmacros_celltypes[8],
                         'not_macrophage')

all_data.combined@meta.data$macros_or_not <- stringr::str_replace_all(all_data.combined@meta.data$macros_or_not ,
                         notmacros_celltypes[9],
                         'not_macrophage')

all_data.combined@meta.data$macros_or_not <- stringr::str_replace_all(all_data.combined@meta.data$macros_or_not ,
                         notmacros_celltypes[10],
                         'not_macrophage')

all_data.combined@meta.data$macros_or_not <- stringr::str_replace_all(all_data.combined@meta.data$macros_or_not ,
                         notmacros_celltypes[11],
                         'not_macrophage')

all_data.combined@meta.data$macros_or_not <- stringr::str_replace_all(all_data.combined@meta.data$macros_or_not ,
                         notmacros_celltypes[12],
                         'not_macrophage')

all_data.combined@meta.data$macros_or_not <- stringr::str_replace_all(all_data.combined@meta.data$macros_or_not ,
                         notmacros_celltypes[13],
                         'not_macrophage')

all_data.combined@meta.data$macros_or_not <- stringr::str_replace_all(all_data.combined@meta.data$macros_or_not ,
                         notmacros_celltypes[14],
                         'not_macrophage')

all_data.combined@meta.data$macros_or_not <- stringr::str_replace_all(all_data.combined@meta.data$macros_or_not ,
                         notmacros_celltypes[15],
                         'not_macrophage')

all_data.combined@meta.data$macros_or_not <- stringr::str_replace_all(all_data.combined@meta.data$macros_or_not ,
                         notmacros_celltypes[16],
                         'not_macrophage')

all_data.combined@meta.data$macros_or_not <- stringr::str_replace_all(all_data.combined@meta.data$macros_or_not ,
                         notmacros_celltypes[17],
                         'not_macrophage')

all_data.combined@meta.data$macros_or_not <- stringr::str_replace_all(all_data.combined@meta.data$macros_or_not ,
                         notmacros_celltypes[18],
                         'not_macrophage')

all_data.combined@meta.data$macros_or_not <- stringr::str_replace_all(all_data.combined@meta.data$macros_or_not ,
                         notmacros_celltypes[19],
                         'not_macrophage')

all_data.combined@meta.data$macros_or_not <- stringr::str_replace_all(all_data.combined@meta.data$macros_or_not ,
                         notmacros_celltypes[20],
                         'not_macrophage')

all_data.combined@meta.data$macros_or_not <- stringr::str_replace_all(all_data.combined@meta.data$macros_or_not ,
                         notmacros_celltypes[21],
                         'not_macrophage')

all_data.combined@meta.data$macros_or_not <- stringr::str_replace_all(all_data.combined@meta.data$macros_or_not ,
                         notmacros_celltypes[22],
                         'not_macrophage')

all_data.combined@meta.data$macros_or_not <- stringr::str_replace_all(all_data.combined@meta.data$macros_or_not ,
                         notmacros_celltypes[23],
                         'not_macrophage')

all_data.combined@meta.data$macros_or_not <- stringr::str_replace_all(all_data.combined@meta.data$macros_or_not ,
                         notmacros_celltypes[24],
                         'not_macrophage')

all_data.combined@meta.data$macros_or_not <- stringr::str_replace_all(all_data.combined@meta.data$macros_or_not ,
                         notmacros_celltypes[25],
                         'not_macrophage')

all_data.combined@meta.data$macros_or_not <- stringr::str_replace_all(all_data.combined@meta.data$macros_or_not ,
                         notmacros_celltypes[26],
                         'not_macrophage')

all_data.combined@meta.data$macros_or_not <- stringr::str_replace_all(all_data.combined@meta.data$macros_or_not ,
                         notmacros_celltypes[27],
                         'not_macrophage')
```

```{r}
Idents(all_data.combined) <- 'seurat_clusters'

p1 <- DotPlot(subset(all_data.combined, subset = (seurat_clusters == '13' | 
                                    seurat_clusters == '16' | 
                                    seurat_clusters == '17' | 
                                    seurat_clusters == '20' | 
                                    seurat_clusters == '23' | 
                                    seurat_clusters == '25' | 
                                    seurat_clusters == '7')),
        features = ram_genes, group.by = 'orig.ident' ) + ggplot2::scale_colour_gradient2(low = "#edf8b1", mid = "#7bccc4", high = "#084081") +
  ggplot2::theme(axis.text.x = ggplot2::element_text(angle = 45, vjust = 1, hjust=1)) + 
  ggplot2::labs(title = "RAM Markers")

p2 <- DotPlot(subset(all_data.combined, subset = (seurat_clusters == '13' | 
                                    seurat_clusters == '16' | 
                                    seurat_clusters == '17' | 
                                    seurat_clusters == '20' | 
                                    seurat_clusters == '23' | 
                                    seurat_clusters == '25' | 
                                    seurat_clusters == '7')),
        features = recam_genes, group.by = 'orig.ident' ) + ggplot2::scale_colour_gradient2(low = "#d4b9da", mid = "#df65b0", high = "#67001f")+
  ggplot2::theme(axis.text.x = ggplot2::element_text(angle = 45, vjust = 1, hjust=1))  + 
  ggplot2::labs(title = "RecAM Markers")

p1+p2
ggplot2::ggsave(here::here('data/all_idents_ram_and_recam_dotplot.png'),
       device = 'png', width = 12, height = 4)
#Saving 16.2 x 9.66 in image  

####
  
Idents(all_data.combined) <- 'seurat_clusters'  
p3 <- DotPlot(subset(all_data.combined, subset = (orig.ident == 'SD7' &
                                                    seurat_clusters == '13' | 
                                    seurat_clusters == '16' | 
                                    seurat_clusters == '17' | 
                                    seurat_clusters == '20' | 
                                    seurat_clusters == '23' | 
                                    seurat_clusters == '25' | 
                                    seurat_clusters == '7')),
        features = ram_genes ) + 
  ggplot2::scale_colour_gradient2(low = "#edf8b1", mid = "#7bccc4", high = "#084081") + 
  ggplot2::theme(axis.text.x = ggplot2::element_text(angle = 45, vjust = 1, hjust=1))+ 
  ggplot2::labs(title = "RAM Markers - SD7 only")

p4 <- DotPlot(subset(all_data.combined, subset = (orig.ident == 'SD7' & 
                                                    seurat_clusters == '13' | 
                                    seurat_clusters == '16' | 
                                    seurat_clusters == '17' | 
                                    seurat_clusters == '20' | 
                                    seurat_clusters == '23' | 
                                    seurat_clusters == '25' | 
                                    seurat_clusters == '7')),
        features = recam_genes) + 
  ggplot2::scale_colour_gradient2(low = "#d4b9da", mid = "#df65b0", high = "#67001f") + 
  ggplot2::theme(axis.text.x = ggplot2::element_text(angle = 45, vjust = 1, hjust=1))+ 
  ggplot2::labs(title = "RecAM Markers - SD7 only")

p3+p4

ggplot2::ggsave(here::here('data/SD7_only_ram_and_recam_dotplot.png'),
       device = 'png', width = 12, height = 4)
#Saving 16.2 x 9.66 in image  
```
```{r}
m2_markers <- c('Alox15', 'Arg1', 'Chil3', 'Chil4', 'Clec7a', 'Egr2', 'Il10', 'Irf4', 'Klf4', 'Mrc1', 'Myc', 'Socs2', 'Tgm2')

m1_markers <- c('Azin1', 'Cd38', 'Cd86', 'Cxcl10', 'Cxcl9', 'Fpr2', 'Gprf8', 'Il12b', 'Il18', 'Il1b', 'Irf5', 'Nfkbiz', 'Nos2', 'Ptgs2', 'Socs3', 'Tlr4', 'Tnf')


m1dp_sd7 <- DotPlot(subset(all_data.combined, subset = (orig.ident == 'SD7' &
                                                    seurat_clusters == '13' | 
                                    seurat_clusters == '16' | 
                                    seurat_clusters == '17' | 
                                    seurat_clusters == '20' | 
                                    seurat_clusters == '23' | 
                                    seurat_clusters == '25' | 
                                    seurat_clusters == '7')),
        features = m1_markers ) + 
  ggplot2::scale_colour_gradient2(low = "#e0ecf4", mid = "#9ebcda", high = "#88419d") + 
  ggplot2::theme(axis.text.x = ggplot2::element_text(angle = 45, vjust = 1, hjust=1))+ 
  ggplot2::labs(title = "M1 markers - SD7 only")

m2dp_sd7 <- DotPlot(subset(all_data.combined, subset = (orig.ident == 'SD7' & 
                                                    seurat_clusters == '13' | 
                                    seurat_clusters == '16' | 
                                    seurat_clusters == '17' | 
                                    seurat_clusters == '20' | 
                                    seurat_clusters == '23' | 
                                    seurat_clusters == '25' | 
                                    seurat_clusters == '7')),
        features = m2_markers) + 
  ggplot2::scale_colour_gradient2(low = "#ffffe5", mid = "#fec44f", high = "#ec7014") + 
  ggplot2::theme(axis.text.x = ggplot2::element_text(angle = 45, vjust = 1, hjust=1))+ 
  ggplot2::labs(title = "M2 Markers - SD7 only")

m1dp_sd7 + m2dp_sd7
ggplot2::ggsave(here::here('data/SD7_only_m1_and_m2_dotplot.png'),
       device = 'png', width = 12, height = 4)


###

Idents(all_data.combined) <- 'seurat_clusters'

m1dp <- DotPlot(subset(all_data.combined, subset = (
                                                    seurat_clusters == '13' | 
                                    seurat_clusters == '16' | 
                                    seurat_clusters == '17' | 
                                    seurat_clusters == '20' | 
                                    seurat_clusters == '23' | 
                                    seurat_clusters == '25' | 
                                    seurat_clusters == '7')),
        features = m1_markers, group.by = 'orig.ident'  ) + 
  ggplot2::scale_colour_gradient2(low = "#e0ecf4", mid = "#9ebcda", high = "#88419d") + 
  ggplot2::theme(axis.text.x = ggplot2::element_text(angle = 45, vjust = 1, hjust=1))+ 
  ggplot2::labs(title = "M1 markers")

m2dp <- DotPlot(subset(all_data.combined, subset = (
                                                    seurat_clusters == '13' | 
                                    seurat_clusters == '16' | 
                                    seurat_clusters == '17' | 
                                    seurat_clusters == '20' | 
                                    seurat_clusters == '23' | 
                                    seurat_clusters == '25' | 
                                    seurat_clusters == '7')),
        features = m2_markers, group.by = 'orig.ident' ) + 
  ggplot2::scale_colour_gradient2(low = "#ffffe5", mid = "#fec44f", high = "#ec7014") + 
  ggplot2::theme(axis.text.x = ggplot2::element_text(angle = 45, vjust = 1, hjust=1))+ 
  ggplot2::labs(title = "M2 Markers")

m1dp+ m2dp
ggplot2::ggsave(here::here('data/all_idents_m1_and_m2_dotplot.png'),
       device = 'png', width = 12, height = 4)
#Saving 16.2 x 9.66 in image  
```

```{r}
#Idents(all_data.combined) <- 'seurat_clusters'
#all_cluster_markers_check <- FindAllMarkers(all_data.combined, only.pos = TRUE, min.pct = 0.25, logfc.threshold = 0.25, assay = 'SCT') #doesn't quite match, probably different Seurat versions...see https://satijalab.org/seurat/articles/announcements.html
```

```{r}


all_cluster_markers_ids <- dplyr::left_join(all_cluster_markers, all_genes_bm, by = c('gene'='mgi_symbol'))
SD7_markers_ids <- dplyr::left_join(SD7_markers, all_genes_bm,  by = c('gene'='mgi_symbol'))


cluster13_markers <- all_cluster_markers_ids %>% dplyr::arrange(desc(avg_log2FC)) %>% dplyr::filter(p_val_adj <= .05) %>% dplyr::filter(cluster == '13')
cluster16_markers <- all_cluster_markers_ids %>% dplyr::arrange(desc(avg_log2FC)) %>% dplyr::filter(p_val_adj <= .05) %>% dplyr::filter(cluster == '16')
cluster17_markers <- all_cluster_markers_ids %>% dplyr::arrange(desc(avg_log2FC)) %>% dplyr::filter(p_val_adj <= .05) %>% dplyr::filter(cluster == '17')
cluster23_markers <- all_cluster_markers_ids %>% dplyr::arrange(desc(avg_log2FC)) %>% dplyr::filter(p_val_adj <= .05) %>% dplyr::filter(cluster == '23')
cluster20_markers <- all_cluster_markers_ids %>% dplyr::arrange(desc(avg_log2FC)) %>% dplyr::filter(p_val_adj <= .05) %>% dplyr::filter(cluster == '20')
cluster25_markers <- all_cluster_markers_ids %>% dplyr::arrange(desc(avg_log2FC)) %>% dplyr::filter(p_val_adj <= .05) %>% dplyr::filter(cluster == '25')
cluster7_markers <- all_cluster_markers_ids %>% dplyr::arrange(desc(avg_log2FC)) %>% dplyr::filter(p_val_adj <= .05) %>% dplyr::filter(cluster == '7')

cluster17_markers <- cluster17_markers %>% dplyr::arrange(desc(avg_log2FC))
openxlsx::write.xlsx(cluster17_markers, file = here::here('data/cluster17_markers.xlsx'))

```

# Table of macrophage subtypes and cluster distribution:
```{r}
cell_types_by_clusters <- table(all_data.combined@meta.data$atlas_predicted_clusters_short, all_data.combined@meta.data$seurat_clusters )

cell_types_by_clusters_SD7_only <- table(
  (subset(all_data.combined, subset = orig.ident == 'SD7'))$atlas_predicted_clusters_short,
  (subset(all_data.combined, subset = orig.ident == 'SD7'))$seurat_clusters)

cell_types_by_clusters_O2D7_only <- table(
  (subset(all_data.combined, subset = orig.ident == 'O2D7'))$atlas_predicted_clusters_short,
  (subset(all_data.combined, subset = orig.ident == 'O2D7'))$seurat_clusters)

cell_types_by_clusters_AirD7_only <- table(
  (subset(all_data.combined, subset = orig.ident == 'AirD7'))$atlas_predicted_clusters_short,
  (subset(all_data.combined, subset = orig.ident == 'AirD7'))$seurat_clusters)


cell_types_by_clusters_O2D60_only <- table(
  (subset(all_data.combined, subset = orig.ident == 'O2D60'))$atlas_predicted_clusters_short,
  (subset(all_data.combined, subset = orig.ident == 'O2D60'))$seurat_clusters)

cell_types_by_clusters_AirD60_only <- table(
  (subset(all_data.combined, subset = orig.ident == 'AirD60'))$atlas_predicted_clusters_short,
  (subset(all_data.combined, subset = orig.ident == 'AirD60'))$seurat_clusters)

openxlsx::write.xlsx(cell_types_by_clusters, file = here::here('data/cell_types_by_clusters.xlsx'))
openxlsx::write.xlsx(cell_types_by_clusters_SD7_only, file = here::here('data/cell_types_by_clusters_SD7_only.xlsx'))
openxlsx::write.xlsx(cell_types_by_clusters_O2D7_only, file = here::here('data/cell_types_by_clusters_O2D7_only.xlsx'))
openxlsx::write.xlsx(cell_types_by_clusters_AirD7_only, file = here::here('data/cell_types_by_clusters_AirD7_only.xlsx'))
openxlsx::write.xlsx(cell_types_by_clusters_O2D60_only, file = here::here('data/cell_types_by_clusters_O2D60_only.xlsx'))
openxlsx::write.xlsx(cell_types_by_clusters_AirD60_only, file = here::here('data/cell_types_by_clusters_AirD60_only.xlsx'))

macrophages_by_clusters_SD7_only <-
  cell_types_by_clusters_SD7_only[rownames(cell_types_by_clusters_SD7_only) %in%
                                    c(
                                      "Alveolar macrophage_Ear2 high",
                                      "Alveolar macrophage_Pclaf high",
                                      "Interstitial macrophage"
                                    ) , ] %>%
  data.frame() %>% data.table::setnames(
    old = c('Var1', 'Var2', 'Freq'),
    new = c('macrophage_type', 'seurat_cluster', 'cell_count')
  )
macrophages_by_clusters_SD7_only$orig_ident <- 'SD7'                                       
                                    
macrophages_by_clusters_O2D7_only <-
  cell_types_by_clusters_O2D7_only[rownames(cell_types_by_clusters_O2D7_only) %in%
                                    c(
                                      "Alveolar macrophage_Ear2 high",
                                      "Alveolar macrophage_Pclaf high",
                                      "Interstitial macrophage"
                                    ) , ] %>%
  data.frame() %>% data.table::setnames(
    old = c('Var1', 'Var2', 'Freq'),
    new = c('macrophage_type', 'seurat_cluster', 'cell_count')
  )
macrophages_by_clusters_O2D7_only$orig_ident <- 'O2D7'   
  

macrophages_by_clusters_AirD7_only <-
  cell_types_by_clusters_AirD7_only[rownames(cell_types_by_clusters_AirD7_only) %in%
                                    c(
                                      "Alveolar macrophage_Ear2 high",
                                      "Alveolar macrophage_Pclaf high",
                                      "Interstitial macrophage"
                                    ) , ] %>%
  data.frame() %>% data.table::setnames(
    old = c('Var1', 'Var2', 'Freq'),
    new = c('macrophage_type', 'seurat_cluster', 'cell_count')
  )
macrophages_by_clusters_AirD7_only$orig_ident <- 'AirD7'   


                                  
macrophages_by_clusters_O2D60_only <-
  cell_types_by_clusters_O2D60_only[rownames(cell_types_by_clusters_O2D60_only) %in%
                                    c(
                                      "Alveolar macrophage_Ear2 high",
                                      "Alveolar macrophage_Pclaf high",
                                      "Interstitial macrophage"
                                    ) , ] %>%
  data.frame() %>% data.table::setnames(
    old = c('Var1', 'Var2', 'Freq'),
    new = c('macrophage_type', 'seurat_cluster', 'cell_count')
  )
macrophages_by_clusters_O2D60_only$orig_ident <- 'O2D60'   
  

macrophages_by_clusters_AirD60_only <-
  cell_types_by_clusters_AirD60_only[rownames(cell_types_by_clusters_AirD60_only) %in%
                                    c(
                                      "Alveolar macrophage_Ear2 high",
                                      "Alveolar macrophage_Pclaf high",
                                      "Interstitial macrophage"
                                    ) , ] %>%
  data.frame() %>% data.table::setnames(
    old = c('Var1', 'Var2', 'Freq'),
    new = c('macrophage_type', 'seurat_cluster', 'cell_count')
  )
macrophages_by_clusters_AirD60_only$orig_ident <- 'AirD60'   
library(dplyr)
all_macrophages_by_clusters_with_idents <- rbind(macrophages_by_clusters_SD7_only, 
      rbind(macrophages_by_clusters_AirD7_only,
            rbind(macrophages_by_clusters_O2D7_only,
                  rbind(macrophages_by_clusters_O2D60_only, macrophages_by_clusters_AirD60_only))))

all_macrophages_by_clusters_with_idents_wide <- all_macrophages_by_clusters_with_idents %>% tidyr::pivot_wider(names_from = orig_ident, values_from = cell_count)


Ear2_macrophages_by_clusters_with_idents_wide <- all_macrophages_by_clusters_with_idents_wide %>% dplyr::filter(macrophage_type == 'Alveolar macrophage_Ear2 high')
Pclaf_macrophages_by_clusters_with_idents_wide <- all_macrophages_by_clusters_with_idents_wide %>% dplyr::filter(macrophage_type == 'Alveolar macrophage_Pclaf high')
Interstitial_macrophages_by_clusters_with_idents_wide <- all_macrophages_by_clusters_with_idents_wide %>% dplyr::filter(macrophage_type == 'Interstitial macrophage')

openxlsx::write.xlsx(Ear2_macrophages_by_clusters_with_idents_wide, file = here::here('data/Ear2_macrophages_by_clusters.xlsx'))
openxlsx::write.xlsx(Pclaf_macrophages_by_clusters_with_idents_wide, file = here::here('data/Pclaf_macrophages_by_clusters.xlsx'))
openxlsx::write.xlsx(Interstitial_macrophages_by_clusters_with_idents_wide, file = here::here('data/Interstitial_macrophages_by_clusters.xlsx'))
```

```{r}
glycolysis_cluster_markers <- dplyr::filter(all_cluster_markers_ids, grepl("R-MMU-70171|R-MMU-70268|R-MMU-71336|R-MMU-71387 ",reactome_ids))

glycolysis_SD7_markers <- dplyr::filter(SD7_markers_ids, grepl("R-MMU-70171|R-MMU-70268|R-MMU-71336|R-MMU-71387 ",reactome_ids))

#R-MMU-70268 is pyruvate metabolism
#R-MMU-70171 is glycolysis
#R-MMU-71336 is pentose phosphate pathway
#R-MMU-71387 is carbohydrate metabolism

#R-MMU-1834941 is STING mediated induction of host immune responses


#Figure 1. Potential sources of error in ECAR measurements. A, homolactic fermentation may result in symport of lactate/H+, making ECAR proportional to substrate-level phosphorylation in glycolysis. B, oxidation of lactate in mitochondria may result in glycolytic flux that is not detectable through ECAR. C, exogenous pyruvate may be reduced to lactate without being directly coupled to glycolytic flux. D, ECAR may be unrelated to lactate/H+ symport (e.g., alternative anion/H+ exchange mechanisms). E, CO2 derived from respiration may induce acidification due to accumulation through bicarbonate and carbonic acid equilibrium. https://www.sciencedirect.com/science/article/pii/S0021925821009418#bib9

glycolysis_cluster_markers %>% dplyr::group_by(cluster)

glycolysis_cluster_markers$gene %>% unique() 
glycolysis_SD7_markers$gene %>% unique()
```

```{r}
cluster_markers_list <- c(cluster13_markers$gene[1:20], 
                          cluster16_markers$gene[1:20], 
                          cluster17_markers$gene[1:20], 
                          cluster20_markers$gene[1:20], 
                          cluster23_markers$gene[1:20],
                          cluster25_markers$gene[1:20], 
                          cluster7_markers$gene[1:20])

cluster_markers_list <- unique(cluster_markers_list)

all_cluster_markers_ids_filtered <- all_cluster_markers_ids[all_cluster_markers_ids$gene %in% cluster_markers_list,]
all_cluster_markers_ids_filtered <- all_cluster_markers_ids_filtered[,c(1:8)]
all_cluster_markers_ids_filtered <- all_cluster_markers_ids_filtered %>% dplyr::filter(p_val_adj <= .05)

#write.table(all_cluster_markers_ids_filtered, file = '/users/jwalla12/dennery_singlecell_reanalysis/data/all_cluster_markers_ids_filtered.csv', row.names = F)
```

```{r}
Idents(sd7_sub) <- 'seurat_clusters'
Idents(all_data.combined) <- 'seurat_clusters'

#clusters are high enough macrophages like 100 total each --  cluster 20 feels like a basal macrophage in that it has few distinguishing marker genes.
clusters_summary <- table(all_data.combined@meta.data$grouped_seurat, all_data.combined@meta.data$orig.ident) %>% rbind()  %>% data.frame() 
clusters_summary <- clusters_summary[ grep('Macrophages', rownames(clusters_summary)),]
clusters_summary[rownames(clusters_summary) %in% c('Macrophages-13', 'Macrophages-16', 'Macrophages-17', 'Macrophages-20', 'Macrophages-23', 'Macrophages-25', 'Macrophages-7'),]


#picked top 20 cluster markers (padj cutoff 0.05, sorted by log2FC)

#Plot the cluster markers in just the Macrophages from all points
cluster_markers_macrophage_clusters_only <-DotPlot(subset(all_data.combined, subset = (atlas_predicted_clusters_grouped == 'Macrophages' & 
                                    seurat_clusters == '13' | 
                                    seurat_clusters == '16' | 
                                    seurat_clusters == '17' | 
                                    seurat_clusters == '20' | 
                                    seurat_clusters == '23' | 
                                    seurat_clusters == '25' | 
                                    seurat_clusters == '7')), assay = 'SCT',features = unique(cluster_markers_list ) ) & ggplot2::coord_flip()& ggplot2::theme(axis.text.x = ggplot2::element_text(angle = 90, vjust = 0.5, hjust=1)) & ggplot2::theme(axis.text=ggplot2::element_text(size=14), legend.text=ggplot2::element_text(size=14), axis.title = ggplot2::element_blank()) & ggplot2::scale_color_gradientn(colors = viridis::turbo(n = 10, direction = 1)) 

#cowplot::save_plot(filename = '/users/jwalla12/dennery_singlecell_reanalysis/data/cluster_markers_macrophage_clusters_only.png', base_width = 7, base_height = 15, plot = cluster_markers_macrophage_clusters_only)

#Plot the cluster markers in just the Macrophages from all points
cluster_markers_macrophage_clusters_only_rotate <-DotPlot(subset(all_data.combined, subset = (atlas_predicted_clusters_grouped == 'Macrophages' & 
                                    seurat_clusters == '13' | 
                                    seurat_clusters == '16' | 
                                    seurat_clusters == '17' | 
                                    seurat_clusters == '20' | 
                                    seurat_clusters == '23' | 
                                    seurat_clusters == '25' | 
                                    seurat_clusters == '7')), assay = 'SCT',features = unique(cluster_markers_list ) ) & ggplot2::theme(axis.text.x = ggplot2::element_text(angle = 90, vjust = 0.5, hjust=1)) & ggplot2::theme(axis.text=ggplot2::element_text(size=14), legend.text=ggplot2::element_text(size=14), axis.title = ggplot2::element_blank()) & ggplot2::scale_color_gradientn(colors = viridis::turbo(n = 10, direction = 1)) 

#cowplot::save_plot(filename = '/users/jwalla12/dennery_singlecell_reanalysis/data/cluster_markers_macrophage_clusters_only_rotate.png', base_width = 18, base_height = 7, plot = cluster_markers_macrophage_clusters_only_rotate)

```

```{r}
glycolysis_cluster_markers$gene %>% unique() 
glycolysis_SD7_markers$gene %>% unique()

glycolysis_cluster_markers_info <- glycolysis_cluster_markers[,c(1:8)]
glycolysis_SD7_markers_info <- glycolysis_SD7_markers[,c(1:8)]

#Plot the cluster glycolysis markers in just the Macrophages from all points
cluster_carbohydrate_markers_macrophage_clusters_only <- DotPlot(subset(all_data.combined, 
                        subset = (atlas_predicted_clusters_grouped == 'Macrophages' & 
                                    seurat_clusters == '13' | 
                                    seurat_clusters == '16' | 
                                    seurat_clusters == '17' | 
                                    seurat_clusters == '20' | 
                                    seurat_clusters == '23' | 
                                    seurat_clusters == '25' | 
                                    seurat_clusters == '7')), assay = 'SCT',features = unique(glycolysis_cluster_markers$gene) ) &ggplot2::coord_flip() & ggplot2::theme(axis.text.x = ggplot2::element_text(angle = 90, vjust = 0.5, hjust=1)) & ggplot2::theme(axis.text=ggplot2::element_text(size=16), legend.text=ggplot2::element_text(size=16), axis.title = ggplot2::element_blank()) & ggplot2::scale_color_gradientn(colors = viridis::turbo(n = 10, direction = 1))   

#cowplot::save_plot(filename = '/users/jwalla12/dennery_singlecell_reanalysis/data/carbohydrate_genes_cluster_markers.png', base_width = 6, base_height = 6, plot = cluster_carbohydrate_markers_macrophage_clusters_only)


#Plot the cluster glycolysis markers in just the Macrophages from all points
cluster_carbohydrate_markers_macrophage_clusters_only_rotate <- DotPlot(subset(all_data.combined, 
                        subset = (atlas_predicted_clusters_grouped == 'Macrophages' & 
                                    seurat_clusters == '13' | 
                                    seurat_clusters == '16' | 
                                    seurat_clusters == '17' | 
                                    seurat_clusters == '20' | 
                                    seurat_clusters == '23' | 
                                    seurat_clusters == '25' | 
                                    seurat_clusters == '7')), assay = 'SCT',features = unique(glycolysis_cluster_markers$gene) ) & ggplot2::theme(axis.text.x = ggplot2::element_text(angle = 90, vjust = 0.5, hjust=1)) & ggplot2::theme(axis.text=ggplot2::element_text(size=16), legend.text=ggplot2::element_text(size=16), axis.title = ggplot2::element_blank()) & ggplot2::scale_color_gradientn(colors = viridis::turbo(n = 10, direction = 1))   

#cowplot::save_plot(filename = '/users/jwalla12/dennery_singlecell_reanalysis/data/carbohydrate_genes_cluster_markers_rotate.png', base_width = 6, base_height = 6, plot = cluster_carbohydrate_markers_macrophage_clusters_only_rotate)

glycolysis_cluster_markers_filtered <- glycolysis_cluster_markers[,c(1:8)]

glycolysis_cluster_markers_filtered <- glycolysis_cluster_markers_filtered %>% dplyr::filter(p_val_adj <= .05)

#write.table(glycolysis_cluster_markers_filtered, file = '/users/jwalla12/dennery_singlecell_reanalysis/data/glycolysis_cluster_markers_filtered.csv', row.names = F)

###
  
#Plot the sd7 glycolysis markers in just the Macrophages from all points

Idents(all_data.combined) <- 'orig.ident'

condition_carbohydrate_markers_macrophage_clusters_only_rotate <- DotPlot(subset(all_data.combined, subset = (atlas_predicted_clusters_grouped == 'Macrophages' & 
                                    seurat_clusters == '13' | 
                                    seurat_clusters == '16' | 
                                    seurat_clusters == '17' | 
                                    seurat_clusters == '20' | 
                                    seurat_clusters == '23' | 
                                    seurat_clusters == '25' | 
                                    seurat_clusters == '7')), assay = 'SCT',features = unique(glycolysis_SD7_markers$gene) ) &ggplot2::coord_flip()& ggplot2::theme(axis.text.x = ggplot2::element_text(angle = 90, vjust = 0.5, hjust=1)) & ggplot2::theme(axis.text=ggplot2::element_text(size=16), legend.text=ggplot2::element_text(size=16), axis.title = ggplot2::element_blank()) & ggplot2::scale_color_gradientn(colors = viridis::turbo(n = 10, direction = 1))    

condition_carbohydrate_markers_macrophage_clusters_only <- DotPlot(subset(all_data.combined, subset = (atlas_predicted_clusters_grouped == 'Macrophages' & 
                                    seurat_clusters == '13' | 
                                    seurat_clusters == '16' | 
                                    seurat_clusters == '17' | 
                                    seurat_clusters == '20' | 
                                    seurat_clusters == '23' | 
                                    seurat_clusters == '25' | 
                                    seurat_clusters == '7')), assay = 'SCT',features = unique(glycolysis_SD7_markers$gene) ) & ggplot2::theme(axis.text.x = ggplot2::element_text(angle = 90, vjust = 0.5, hjust=1)) & ggplot2::theme(axis.text=ggplot2::element_text(size=16), legend.text=ggplot2::element_text(size=16), axis.title = ggplot2::element_blank()) & ggplot2::scale_color_gradientn(colors = viridis::turbo(n = 10, direction = 1))    


#cowplot::save_plot(filename = '/users/jwalla12/dennery_singlecell_reanalysis/data/condition_carbohydrate_markers_macrophage_clusters_only.png', base_width = 6, base_height = 6, plot = condition_carbohydrate_markers_macrophage_clusters_only)

#cowplot::save_plot(filename = '/users/jwalla12/dennery_singlecell_reanalysis/data/condition_carbohydrate_markers_macrophage_clusters_only_rotate.png', base_width = 6, base_height = 6, plot = condition_carbohydrate_markers_macrophage_clusters_only_rotate)



glycolysis_SD7_markers <- glycolysis_SD7_markers[,c(1:8)]

#write.table(glycolysis_SD7_markers , file = '/users/jwalla12/dennery_singlecell_reanalysis/data/glycolysis_SD7_markers_filtered.csv', row.names = F)

```

```{r}
#gapdh featureplot  
Idents(all_data.combined) <- 'seurat_clusters'

fp1 <- FeaturePlot(subset(all_data.combined, 
                   subset = (atlas_predicted_clusters_grouped == 'Macrophages' & 
                                    seurat_clusters == '13' | 
                                    seurat_clusters == '16' | 
                                    seurat_clusters == '17' | 
                                    seurat_clusters == '20' | 
                                    seurat_clusters == '23' | 
                                    seurat_clusters == '25' | 
                                    seurat_clusters == '7')), 
                   features = c("Gapdh"), order = T, slot = 'scale.data', label = TRUE, repel = TRUE, pt.size =1, split.by = 'orig.ident', keep.scale = 'feature', reduction = 'tsne') & 
              ggplot2::scale_color_gradientn(colors = rev(RColorBrewer::brewer.pal(10, 'Spectral'))[1:9]) & ggplot2::labs(color = "SCTransform scale data")

dp1 <- DimPlot(subset(all_data.combined, 
                   subset = (atlas_predicted_clusters_grouped == 'Macrophages' & 
                                    seurat_clusters == '13' | 
                                    seurat_clusters == '16' | 
                                    seurat_clusters == '17' | 
                                    seurat_clusters == '20' | 
                                    seurat_clusters == '23' | 
                                    seurat_clusters == '25' | 
                                    seurat_clusters == '7')), reduction = 'tsne', label = T, label.box = T)
fp1+dp1 + patchwork::plot_layout(ncol=3, nrow=2)
#save.image('/users/jwalla12/dennery_singlecell_reanalysis/analysis.RData')
```

```{r}
#load('/users/jwalla12/dennery_singlecell_reanalysis/analysis.RData')
```

#July 15 2024

```{r}


grep('itih', rownames(all_data.combined), ignore.case=T, value=T) #Detected Itih1, 2, 3, 4, and 5.

grep('itih', all_cluster_markers_ids$gene, ignore.case=T, value=T) #Itih4 and 5 were markers..

dplyr::filter(all_cluster_markers_ids, gene == 'Itih1' | gene == 'Itih2' | gene == 'Itih3' | gene == 'Itih4' | gene == 'Itih5') #Itih4 is marker for Cluster3, Itih5 is marker for clusters 18 and 33

DotPlot(all_data.combined, assay = 'SCT',features = c('Itih4','Itih5')) & ggplot2::theme(axis.text.x = ggplot2::element_text(angle = 90, vjust = 0.5, hjust=1)) & ggplot2::theme(axis.text=ggplot2::element_text(size=16), legend.text=ggplot2::element_text(size=16), axis.title = ggplot2::element_blank()) & ggplot2::scale_color_gradientn(colors = viridis::turbo(n = 10, direction = 1))    
```

```{r}
library(dplyr)
clusters_vs_idents <- as.data.frame(table(all_data.combined@meta.data$seurat_clusters, all_data.combined@meta.data$orig.ident)) %>% tidyr::pivot_wider(names_from = Var2, values_from = Freq)
colnames(clusters_vs_idents) <- c('Seurat_clusters','AirD7','O2D7','SD7','AirD60','O2D60')

clusters_vs_shorttypes <- as.data.frame(table(all_data.combined@meta.data$seurat_clusters, all_data.combined@meta.data$atlas_predicted_clusters_short)) %>% tidyr::pivot_wider(names_from = Var2, values_from = Freq) %>% data.frame()
rownames(clusters_vs_shorttypes) <- clusters_vs_shorttypes$Var1
clusters_vs_shorttypes$Var1 <- NULL
clusters_vs_shorttypes$Seurat_clusters <- rownames(clusters_vs_shorttypes)
clusters_vs_shorttypes <- clusters_vs_shorttypes %>% dplyr::select('Seurat_clusters', colnames(clusters_vs_shorttypes)[1:30])

clusters_vs_groupedtypes <- as.data.frame(table(all_data.combined@meta.data$seurat_clusters, all_data.combined@meta.data$atlas_predicted_clusters_grouped)) %>% tidyr::pivot_wider(names_from = Var2, values_from = Freq) %>% data.frame()
rownames(clusters_vs_groupedtypes) <- clusters_vs_groupedtypes$Var1
clusters_vs_groupedtypes$Var1 <- NULL
clusters_vs_groupedtypes$Seurat_clusters <- rownames(clusters_vs_groupedtypes)
clusters_vs_groupedtypes <- clusters_vs_groupedtypes %>% dplyr::select('Seurat_clusters', colnames(clusters_vs_groupedtypes)[1:7])

idents_vs_shorttypes <- as.data.frame(table(all_data.combined@meta.data$atlas_predicted_clusters_short, all_data.combined@meta.data$orig.ident)) %>% tidyr::pivot_wider(names_from = Var2, values_from = Freq) %>% data.frame()
idents_vs_shorttypes$cell_types_short <- idents_vs_shorttypes$Var1
idents_vs_shorttypes$Var1 <- NULL

idents_vs_shorttypes <- idents_vs_shorttypes %>% dplyr::select('cell_types_short', colnames(idents_vs_shorttypes)[1:5])


idents_vs_groupedtypes <- as.data.frame(table(all_data.combined@meta.data$atlas_predicted_clusters_grouped, all_data.combined@meta.data$orig.ident)) %>% tidyr::pivot_wider(names_from = Var2, values_from = Freq) %>% data.frame()
idents_vs_groupedtypes$cell_types_grouped <- idents_vs_groupedtypes$Var1
idents_vs_groupedtypes$Var1 <- NULL

idents_vs_groupedtypes <- idents_vs_groupedtypes %>% dplyr::select('cell_types_grouped', colnames(idents_vs_groupedtypes)[1:5])



#openxlsx::write.xlsx(clusters_vs_idents, file = here::here('clusters_vs_idents.xlsx'))

#openxlsx::write.xlsx(clusters_vs_shorttypes, file = here::here('clusters_vs_shorttypes.xlsx'))

#openxlsx::write.xlsx(clusters_vs_groupedtypes, file = here::here('clusters_vs_groupedtypes.xlsx'))

#openxlsx::write.xlsx(idents_vs_shorttypes, file = here::here('idents_vs_shorttypes.xlsx'))

#openxlsx::write.xlsx(idents_vs_groupedtypes, file = here::here('idents_vs_groupedtypes.xlsx'))


#save.image('/oscar/data/cbc/Dennery_scRNAseq/dennery_singlecell_reanalysis/analysis.RData')

```



```{r}
table(all_data.combined@meta.data$seurat_clusters, all_data.combined@meta.data$atlas_predicted_clusters_short)
```

```{r}
table(all_data.combined@meta.data$seurat_clusters,all_data.combined@meta.data$atlas_predicted_clusters_grouped) #Cluster 3 has very few senescent cells (11), 18 only has 1, 33 has 0.
```

```{r}
table(all_data.combined@meta.data$atlas_predicted_clusters_short,all_data.combined@meta.data$orig.ident) #Cluster 3 has very few senescent cells (11), 18 only has 1, 33 has 0.
```

```{r}
dplyr::filter(all_cluster_markers_ids, cluster == '17') %>% dplyr::arrange(desc(avg_log2FC))
```

What about monocyte genes? Could use the monocyte marker containing clusters as the root of the trajectory.
Look in emails -- Hongwei sent me a list of genes.
```{r}
monocyte_genes <- c(grep('Ecm1', rownames(all_data.combined), ignore.case=T, value=T),
  grep('Fabp4', rownames(all_data.combined), ignore.case=T, value=T),
  grep('Spp1', rownames(all_data.combined), ignore.case=T, value=T),
  grep('Runx2', rownames(all_data.combined), ignore.case=T, value=T),
  grep('Ly6c', rownames(all_data.combined), ignore.case=T, value=T),
  grep('Irf', rownames(all_data.combined), ignore.case=T, value=T))


all_monocytes <- all_cluster_markers_ids[all_cluster_markers_ids$gene %in% monocyte_genes,] 

all_monocytes %>% dplyr::filter( p_val_adj <= 1e-5) %>% dplyr::filter(avg_log2FC >= .3)
#cluster 6, 15, and 37 all express some of the monocyte markers and are all in one chunk of the umap, should be able to select them as root...

```

#Run monocle3 here using the monocyte-like clusters as root
```{r}
#library(monocle3)
#https://stuartlab.org/signac/articles/monocle.html#building-trajectories-with-monocle-3
#https://github.com/satijalab/seurat-wrappers/blob/master/docs/monocle3.md
#https://github.com/cole-trapnell-lab/monocle3/issues/328


all_data.combined.cds <- SeuratWrappers::as.cell_data_set(all_data.combined)


# find all possible partitions
all_partitions <- unique(all_data.combined.cds@clusters$UMAP$partitions)
all_partitions <- all_partitions[all_partitions != "1"]

# set all partitions to 1
all_data.combined.cds@clusters$UMAP$partitions[all_data.combined.cds@clusters$UMAP$partitions %in% all_partitions] <- "1"



all_data.combined.cds <- monocle3::cluster_cells(cds = all_data.combined.cds, reduction_method = "UMAP")
all_data.combined.cds <- monocle3::learn_graph(all_data.combined.cds, use_partition = F)
#monocle3::plot_cells(all_data.combined.cds)


#cell_ids <- which(SummarizedExperiment::colData(all_data.combined.cds)[, "ident"] == 6|15|37)
#closest_vertex <- all_data.combined.cds@principal_graph_aux[["UMAP"]]$pr_graph_cell_proj_closest_vertex
#closest_vertex <- as.matrix(closest_vertex[colnames(all_data.combined.cds), ])

#SummarizedExperiment::colData(all_data.combined.cds)

cell_ids_6 <- which(SummarizedExperiment::colData(all_data.combined.cds)[, "seurat_clusters"] == 6)
cell_ids_15 <- which(SummarizedExperiment::colData(all_data.combined.cds)[, "seurat_clusters"] == 15)
cell_ids_37 <- which(SummarizedExperiment::colData(all_data.combined.cds)[, "seurat_clusters"] == 37)
cell_ids <- c(cell_ids_6, cell_ids_15, cell_ids_37)

cell_ids <- rownames(SummarizedExperiment::colData(all_data.combined.cds)[c(cell_ids),])
#cluster 6, 15, and 37 all express some of the monocyte markers and are all in one chunk of the umap, should be able to select them as root...

all_data.combined.cds <- monocle3::order_cells(all_data.combined.cds, root_cells = cell_ids)

cell_pseudotimes <- monocle3::pseudotime(all_data.combined.cds, reduction_method = "UMAP")

all_data.combined <- AddMetaData(all_data.combined, cell_pseudotimes, col.name = 'pseudotime')




pt_fp <- FeaturePlot(all_data.combined,
                            features = "pseudotime", 
                            order = T, 
                            pt.size = 1) & 
              ggplot2::scale_color_gradientn(
                colors = 
                  viridis::turbo(n = 10, direction = 1)) &
              ggplot2::theme(legend.position = 'right')

#13,16,17,20,23,25,7
Idents(all_data.combined) <- 'atlas_predicted_clusters_short'
apcs_fp <- DimPlot(all_data.combined, label = F)

Idents(all_data.combined) <- 'atlas_predicted_clusters_grouped'
apcg_fp <- DimPlot(all_data.combined, label = F)
Idents(all_data.combined) <- 'atlas_predicted_clusters_short'


ear2 <- FeaturePlot(subset(all_data.combined, 
                   subset = (atlas_predicted_clusters_short == "Alveolar macrophage_Ear2 high")),
                            features = "pseudotime",
                            order = T, 
                            pt.size = 1, label =F) & 
              ggplot2::scale_color_gradientn(
                colors = 
                  viridis::turbo(n = 10, direction = 1), limits = c(0,50)) &
              ggplot2::theme(legend.position = 'right') & ggplot2::labs(title = 'Alveolar macrophage_Ear2 high')



pclaf <- FeaturePlot(subset(all_data.combined, 
                   subset = ( atlas_predicted_clusters_short == "Alveolar macrophage_Pclaf high" )),
                            features = "pseudotime",
                            order = T, 
                            pt.size = 1, label =F) & 
              ggplot2::scale_color_gradientn(
                colors = 
                  viridis::turbo(n = 10, direction = 1), limits = c(0,50)) &
              ggplot2::theme(legend.position = 'right') & ggplot2::labs(title = 'Alveolar macrophage_Pclaf high')

inters <- FeaturePlot(subset(all_data.combined, 
                   subset = (atlas_predicted_clusters_short == "Interstitial macrophage")),
                            features = "pseudotime",
                            order = T, 
                            pt.size = 1, label =F) & 
              ggplot2::scale_color_gradientn(
                colors = 
                  viridis::turbo(n = 10, direction = 1), limits = c(0,50)) &
              ggplot2::theme(legend.position = 'right') & ggplot2::labs(title = 'Interstitial macrophage')
                   
pt_fp_macros <- ear2/pclaf/inters                 
                   

colors <- grDevices::colors()[grep('gr(a|e)y|light', grDevices::colors(), invert = T)]
cluster_colors <- sample(x = colors, size = all_data.combined@meta.data$atlas_predicted_clusters_grouped %>% unique() %>% length())

library(colorspace)

hcl_colors <- hcl.colors(all_data.combined@meta.data$atlas_predicted_clusters_grouped %>% unique() %>% length(), "Set 2")
names(hcl_colors) <- all_data.combined@meta.data$atlas_predicted_clusters_grouped %>% unique() 


apcg_dimplot <- DimPlot(
  all_data.combined,
  group.by = "atlas_predicted_clusters_grouped",
  cols = hcl_colors) 
apcg_dimplot

Idents(all_data.combined) <- 'atlas_predicted_clusters_short'
apcs_fp <- DimPlot(all_data.combined)


Idents(all_data.combined) <- 'seurat_clusters'
sc_fp <- DimPlot(all_data.combined, label = T) + NoLegend()

plot <- apcg_fp  /pt_fp 
ggplot2::ggsave('/oscar/data/cbc/Dennery_scRNAseq/dennery_singlecell_reanalysis/plot.png', plot)

plot2 <- sc_fp /pt_fp 
ggplot2::ggsave('/oscar/data/cbc/Dennery_scRNAseq/dennery_singlecell_reanalysis/plot2.png', plot2, height = 9, width = 7)

plot3 <- apcg_dimplot / pt_fp 
ggplot2::ggsave('/oscar/data/cbc/Dennery_scRNAseq/dennery_singlecell_reanalysis/plot3.png', plot3, height = 9, width = 7)


ggplot2::ggsave('/oscar/data/cbc/Dennery_scRNAseq/dennery_singlecell_reanalysis/macros_plot.png',pt_fp_macros, height = 9, width = 4)

save.image('/oscar/data/cbc/Dennery_scRNAseq/dennery_singlecell_reanalysis/analysis.RData')

```

```{r}
all_data.combined


monocyte_markers_in_macros <- all_cluster_markers_ids[all_cluster_markers_ids$gene %in% monocyte_genes,] %>% dplyr::filter(cluster == 13|cluster ==  16|cluster == 17|cluster == 20|cluster == 23|cluster == 25|cluster == 7)

monocyte_markers_in_macros %>% dplyr::filter(cluster == '17') #cluster 17 up regulates Spp1
monocyte_markers_in_macros %>% dplyr::filter(cluster == '13') #cluster 13 up regulates Spp1, Runx2, Irf2bp2
monocyte_markers_in_macros %>% dplyr::filter(cluster == '16') #cluster 16 nothing 
monocyte_markers_in_macros %>% dplyr::filter(cluster == '20') #cluster 16 down regulates Irf2bp2
monocyte_markers_in_macros %>% dplyr::filter(cluster == '23') #cluster 23 up regulates Irf8, Irf5, and Irf2bp2
monocyte_markers_in_macros %>% dplyr::filter(cluster == '25') #cluster 25 up regulates Spp1, Fabp4, Runx2, and Irf2bp2

monocyte_de_markers <- c('Spp1','Runx2','Irf2bp2','Irf8','Irf5','Fabp4')

monocyte_markers_in_macros_only_rotate <- DotPlot(subset(all_data.combined, subset = (atlas_predicted_clusters_grouped == 'Macrophages' & 
                                    seurat_clusters == '13' | 
                                    seurat_clusters == '16' | 
                                    seurat_clusters == '17' | 
                                    seurat_clusters == '20' | 
                                    seurat_clusters == '23' | 
                                    seurat_clusters == '25' | 
                                    seurat_clusters == '7')), assay = 'SCT',features = monocyte_de_markers ) &ggplot2::coord_flip()& ggplot2::theme(axis.text.x = ggplot2::element_text(angle = 90, vjust = 0.5, hjust=1)) & ggplot2::theme(axis.text=ggplot2::element_text(size=16), legend.text=ggplot2::element_text(size=16), axis.title = ggplot2::element_blank()) & ggplot2::scale_color_gradientn(colors = viridis::turbo(n = 10, direction = 1))    

monocyte_markers_in_macros_only <- DotPlot(subset(all_data.combined, subset = (atlas_predicted_clusters_grouped == 'Macrophages' & 
                                    seurat_clusters == '13' | 
                                    seurat_clusters == '16' | 
                                    seurat_clusters == '17' | 
                                    seurat_clusters == '20' | 
                                    seurat_clusters == '23' | 
                                    seurat_clusters == '25' | 
                                    seurat_clusters == '7')), assay = 'SCT',features = monocyte_de_markers ) & ggplot2::theme(axis.text.x = ggplot2::element_text(angle = 90, vjust = 0.5, hjust=1)) & ggplot2::theme(axis.text=ggplot2::element_text(size=16), legend.text=ggplot2::element_text(size=16), axis.title = ggplot2::element_blank()) & ggplot2::scale_color_gradientn(colors = viridis::turbo(n = 10, direction = 1))  

```

```{r}
library(msigdbr)

h_gene_sets <- msigdbr(species = "mouse", category = "H")

oxphos_genes <- h_gene_sets %>% dplyr::filter(gs_name == 'HALLMARK_OXIDATIVE_PHOSPHORYLATION')

oxphos_genes <- oxphos_genes$gene_symbol %>% unique()

oxphos_genes_in_macros <- all_cluster_markers_ids[all_cluster_markers_ids$gene %in% oxphos_genes,] %>% dplyr::filter(cluster == 13|cluster ==  16|cluster == 17|cluster == 20|cluster == 23|cluster == 25|cluster == 7)

#no changes in the oxphos in the Air/O2 data - they use the seahorse for this. you add substrate over intervals of time and gives you a plot of the cell o2 consumption rate (some colorimetric read out probably).
#any changes in the oxphos genes in cluster 17 across time points also? if we compare all the D7 time points do we see changes in the differences 

oxphos_genes_in_macros %>% dplyr::filter(cluster == '17') %>% dplyr::filter(avg_log2FC >0) %>% nrow() #cluster 17 up regulates 51 oxphos genes
oxphos_genes_in_macros %>% dplyr::filter(cluster == '13') %>% dplyr::filter(avg_log2FC >0) %>% nrow() #cluster 13 up regulates 43 oxphos genes
oxphos_genes_in_macros %>% dplyr::filter(cluster == '16') %>% dplyr::filter(avg_log2FC >0) %>% nrow() #cluster 16 up regulates 3 oxphos genes
oxphos_genes_in_macros %>% dplyr::filter(cluster == '20') %>% dplyr::filter(avg_log2FC >0) %>% nrow() #cluster 16 up regulates 18 oxphos genes
oxphos_genes_in_macros %>% dplyr::filter(cluster == '23') %>% dplyr::filter(avg_log2FC >0) %>% nrow() #cluster 23 up regulates 4 oxphos genes
oxphos_genes_in_macros %>% dplyr::filter(cluster == '25') %>% dplyr::filter(avg_log2FC >0) %>% nrow() #cluster 25 up regulates 44 oxphos genes


oxphos_de_markers <- c(oxphos_genes_in_macros %>% dplyr::filter(cluster == '17') %>% dplyr::filter(p_val_adj <= .05) %>% dplyr::filter(avg_log2FC >0) %>% dplyr::select('gene') %>% unlist(),
                       oxphos_genes_in_macros %>% dplyr::filter(cluster == '13') %>% dplyr::filter(p_val_adj <= .05) %>% dplyr::filter(avg_log2FC >0) %>% dplyr::select('gene') %>% unlist(),
                       oxphos_genes_in_macros %>% dplyr::filter(cluster == '16') %>% dplyr::filter(p_val_adj <= .05) %>% dplyr::filter(avg_log2FC >0) %>% dplyr::select('gene') %>% unlist(),
                       oxphos_genes_in_macros %>% dplyr::filter(cluster == '20') %>% dplyr::filter(p_val_adj <= .05) %>% dplyr::filter(avg_log2FC >0) %>% dplyr::select('gene') %>% unlist(),
                       oxphos_genes_in_macros %>% dplyr::filter(cluster == '23') %>% dplyr::filter(p_val_adj <= .05) %>% dplyr::filter(avg_log2FC >0) %>% dplyr::select('gene') %>% unlist(),
                       oxphos_genes_in_macros %>% dplyr::filter(cluster == '25') %>% dplyr::filter(p_val_adj <= .05) %>% dplyr::filter(avg_log2FC >0) %>% dplyr::select('gene') %>% unlist()) %>% unique()

oxphos_17_de_markers <- c(oxphos_genes_in_macros %>% dplyr::filter(cluster == '17') %>% dplyr::filter(p_val_adj <= .05) %>% dplyr::filter(avg_log2FC >0) %>% dplyr::select('gene') %>% unlist())

names(oxphos_17_de_markers) <- NULL


oxphos_markers_in_macros_only_rotate <- DotPlot(subset(all_data.combined, subset = (atlas_predicted_clusters_grouped == 'Macrophages' & 
                                    seurat_clusters == '13' | 
                                    seurat_clusters == '16' | 
                                    seurat_clusters == '17' | 
                                    seurat_clusters == '20' | 
                                    seurat_clusters == '23' | 
                                    seurat_clusters == '25' | 
                                    seurat_clusters == '7')), assay = 'SCT',features = oxphos_de_markers ) &ggplot2::coord_flip()& ggplot2::theme(axis.text.x = ggplot2::element_text(angle = 90, vjust = 0.5, hjust=1)) & ggplot2::theme(axis.text=ggplot2::element_text(size=12), legend.text=ggplot2::element_text(size=12), axis.title = ggplot2::element_blank()) & ggplot2::scale_color_gradientn(colors = viridis::turbo(n = 10, direction = 1))    

oxphos_markers_in_macros_only <- DotPlot(subset(all_data.combined, subset = (atlas_predicted_clusters_grouped == 'Macrophages' & 
                                    seurat_clusters == '13' | 
                                    seurat_clusters == '16' | 
                                    seurat_clusters == '17' | 
                                    seurat_clusters == '20' | 
                                    seurat_clusters == '23' | 
                                    seurat_clusters == '25' | 
                                    seurat_clusters == '7')), assay = 'SCT',features = oxphos_de_markers ) & ggplot2::theme(axis.text.x = ggplot2::element_text(angle = 90, vjust = 0.5, hjust=1)) & ggplot2::theme(axis.text=ggplot2::element_text(size=16), legend.text=ggplot2::element_text(size=16), axis.title = ggplot2::element_blank()) & ggplot2::scale_color_gradientn(colors = viridis::turbo(n = 10, direction = 1))  




oxphos_markers_in_17_only_rotate <- DotPlot(subset(all_data.combined, subset = (atlas_predicted_clusters_grouped == 'Macrophages' & 
                                    seurat_clusters == '13' | 
                                    seurat_clusters == '16' | 
                                    seurat_clusters == '17' | 
                                    seurat_clusters == '20' | 
                                    seurat_clusters == '23' | 
                                    seurat_clusters == '25' | 
                                    seurat_clusters == '7')), assay = 'SCT',features = oxphos_17_de_markers ) &ggplot2::coord_flip()& ggplot2::theme(axis.text.x = ggplot2::element_text(angle = 90, vjust = 0.5, hjust=1)) & ggplot2::theme(axis.text=ggplot2::element_text(size=16), legend.text=ggplot2::element_text(size=16), axis.title = ggplot2::element_blank()) & ggplot2::scale_color_gradientn(colors = viridis::turbo(n = 10, direction = 1))    

oxphos_markers_in_17_only <- DotPlot(subset(all_data.combined, subset = (atlas_predicted_clusters_grouped == 'Macrophages' & 
                                    seurat_clusters == '13' | 
                                    seurat_clusters == '16' | 
                                    seurat_clusters == '17' | 
                                    seurat_clusters == '20' | 
                                    seurat_clusters == '23' | 
                                    seurat_clusters == '25' | 
                                    seurat_clusters == '7')), assay = 'SCT',features = oxphos_17_de_markers ) & ggplot2::theme(axis.text.x = ggplot2::element_text(angle = 90, vjust = 0.5, hjust=1)) & ggplot2::theme(axis.text=ggplot2::element_text(size=16), legend.text=ggplot2::element_text(size=16), axis.title = ggplot2::element_blank()) & ggplot2::scale_color_gradientn(colors = viridis::turbo(n = 10, direction = 1))  



Idents(all_data.combined) <- 'orig.ident'

oxphos_markers_in_origidents<- DotPlot(subset(all_data.combined, subset = (atlas_predicted_clusters_grouped == 'Macrophages' & 
                                    seurat_clusters == '13' | 
                                    seurat_clusters == '16' | 
                                    seurat_clusters == '17' | 
                                    seurat_clusters == '20' | 
                                    seurat_clusters == '23' | 
                                    seurat_clusters == '25' | 
                                    seurat_clusters == '7')), assay = 'SCT',features = oxphos_17_de_markers, group.by = 'orig.ident' ) & ggplot2::theme(axis.text.x = ggplot2::element_text(angle = 90, vjust = 0.5, hjust=1)) & ggplot2::theme(axis.text=ggplot2::element_text(size=16), legend.text=ggplot2::element_text(size=16), axis.title = ggplot2::element_blank()) & ggplot2::scale_color_gradientn(colors = viridis::turbo(n = 10, direction = 1))  

ggplot2::ggsave('/oscar/data/cbc/Dennery_scRNAseq/dennery_singlecell_reanalysis/oxphos_markers_in_origidents.png',oxphos_markers_in_origidents, width = 15.6, height = 10.2)


oxphos_markers_in_origidents_unfiltered<- DotPlot(all_data.combined, assay = 'SCT',features = oxphos_17_de_markers, group.by = 'orig.ident' ) & ggplot2::theme(axis.text.x = ggplot2::element_text(angle = 90, vjust = 0.5, hjust=1)) & ggplot2::theme(axis.text=ggplot2::element_text(size=16), legend.text=ggplot2::element_text(size=16), axis.title = ggplot2::element_blank()) & ggplot2::scale_color_gradientn(colors = viridis::turbo(n = 10, direction = 1))  

ggplot2::ggsave('/oscar/data/cbc/Dennery_scRNAseq/dennery_singlecell_reanalysis/oxphos_markers_in_origidents_unfiltered.png',oxphos_markers_in_origidents_unfiltered,  width = 15.6, height = 10.2)

```
#July 15 2024 end
